\chapter{Software Testing of Maritime Surveillance and Trajectory Prediction}

This chapter presents the rigorous testing methodology employed to validate the multi-stage surveillance pipeline. It discusses unit, integration, and system-level testing, detailing the controlled experimental setup, stress testing under high computational loads, and specific functional test cases designed to ensure robust performance.

\section{Introduction to the Testing Methodology}

Validating a complex, multi-stage computer vision and trajectory prediction system requires a rigorous, multi-tiered testing strategy. Because the proposed architecture relies on the integration of distinct functional modules, ranging from optical ship detection and appearance-based tracking to geographic sensor fusion and recurrent neural networks, evaluating the software demanded more than just a superficial review of the final output. The testing phase was specifically structured to isolate individual components before assessing their collective performance in a simulated operational environment. This systematic approach ensures that localized logical faults or mathematical anomalies within a single subsystem do not cascade into larger failures during real-time surveillance monitoring.

\section{Unit and Integration Testing}

Prior to assembling the overarching surveillance pipeline, each core software module was subjected to targeted unit testing. For the vision-based components, this involved passing static maritime imagery through the YOLOv8 detector to verify that bounding box coordinates were being extracted and formatted correctly according to the strict structural expectations of the DeepOCSORT tracking algorithm. Concurrently, the mathematical utility functions responsible for calculating Euclidean distances, deriving inter-frame velocity, and executing the Haversine nearest-neighbor algorithm were tested against manually calculated control datasets. This guaranteed that the fundamental geometric and trigonometric operations underlying the system's tracking logic functioned with absolute numerical precision.

Following baseline validation, the evaluation focus shifted toward integration testing to scrutinize the data interfaces between the asynchronously operating modules. A critical integration test involved bridging the continuous computer vision pipeline with the geographic sensor fusion engine. The system was challenged to reliably associate noisy, high-frequency optical coordinate updates with sparse, structured AIS telemetry broadcasts. By simulating scenarios with delayed or incomplete AIS packets, the testing verified that the Kalman filter correctly bridged these telemetry gaps, successfully stabilizing the fused position state without crashing the parent process or disrupting the active video rendering.

\section{System Testing}
\subsection{Experimental Setup}

To comprehensively evaluate the end-to-end performance of the integrated pipeline under realistic operational constraints, all system testing was conducted within a rigidly controlled hardware and software environment. Maintaining a standardized baseline was essential for guaranteeing the reproducibility of all quantitative results. 

Handling simultaneous video ingestion, multi-target deep learning inference, and recurrent trajectory forecasting demands significant computational bandwidth. As a result, the implementation relied heavily on parallel GPU acceleration to maintain a fluid operational state. Developing and evaluating the software across mixed operating system environments ensured that the dependencies remained highly adaptable and were not strictly bound to a single ecosystem. 

Furthermore, memory allocation dynamics were closely monitored during the setup phase. The PyTorch tensor allocations required careful optimization to ensure that high-resolution video frames did not overflow the available VRAM. By systematically restricting batch sizes during live inference, the framework was successfully stabilized within standard commercial hardware limitations.

Beyond hardware constraints, the networking environment was also strictly regulated to simulate real-world port conditions. Since the system depends on continuous WebSocket streams to broadcast trajectory predictions to the operator dashboard, evaluating network latency was paramount. A localized gigabit switch was employed to route the telemetry data between the inference backend and the frontend visualization clients. This closed-loop network topology intentionally removed external internet volatility from the testing equation, ensuring that any measured latency could be directly attributed to algorithmic processing rather than fluctuating bandwidth.

Additionally, data storage and retrieval mechanisms were scrutinized during this setup phase. The system requires rapid reading of high-definition video files for offline simulation and fast writing capabilities for logging historical tracking tensors. Solid-state drives (SSDs) utilizing the NVMe protocol were exclusively used to prevent I/O bottlenecks. If traditional mechanical hard drives had been employed, the read/write latency would have severely hampered the YOLOv8 model's ability to ingest frames at the required 30 frames per second, thus skewing the entire performance evaluation. Table~\ref{c5:tab:hardware} details the specific architectural specifications and core software libraries deployed throughout the testing phase.

\begin{table}[H]
\centering
\caption{Hardware and Software Configuration Used for Experimental Evaluation}
\label{c5:tab:hardware}
\begin{tabular}{|p{0.34\textwidth}|p{0.34\textwidth}|}
\hline
\textbf{Component} & \textbf{Specification}\\
\hline
Operating System & Windows 11 / Ubuntu 22.04\\
\hline
Processor & Intel Core i5 12th gen / AMD Ryzen 7\\
\hline
\gls{gpu} & \gls{nvidia} \gls{gpu} with \gls{cuda} support\\
\hline
\gls{ram} & 8 GB\\
\hline
Python Version & 3.10\\
\hline
\gls{pytorch} Version & 2.2.0\\
\hline
\gls{cuda} Version & 11.8\\
\hline
\gls{opencv} Version & 4.9.0\\
\hline
Ultralytics \gls{yolov8} & v8.0\\
\hline
\gls{deepocsort} (boxmot) & v10.0\\
\hline
Evaluation Dataset & \gls{smd} v1.0\\
\hline
\end{tabular}
\end{table}

\subsection{Standardization of System Parameters}

Beyond establishing the physical hardware constraints, preserving strict control over the internal algorithmic parameters was vital for isolating the true performance metrics of the embedded models. Even minor variations in tracking thresholds or neural network configurations can drastically alter the final output, introducing uncontrolled bias into the evaluation process. To prevent such discrepancies, a static set of operational parameters was locked down across all experimental runs. 

These rigid configurations dictated everything from the minimum statistical confidence required to register a valid vessel detection, to the exact sequence length ingested by the predictive memory networks. Furthermore, the geographic boundaries governing the optical-to-AIS matching algorithms were explicitly fixed to ensure consistency when assessing sensor fusion accuracy. 

Establishing these constants prevents the phenomenon of hyperparameter drift, where unrecorded tweaks during testing make results irreproducible. For instance, the track minimum hit count and maximum age parameters fundamentally alter the sensitivity of the DeepOCSORT algorithm; changing them mid-evaluation would completely invalidate any comparative tracking analysis. Table~\ref{c5:tab:parameters} lists the critical thresholds and tracking parameters actively enforced throughout the system-level testing.

\begin{table}[H]
\centering
\caption{System Parameters Used During Experimental Evaluation}
\label{c5:tab:parameters}
\begin{tabular}{|p{0.34\textwidth}|p{0.34\textwidth}|}
\hline
\textbf{Parameter} & \textbf{Value}\\
\hline
Detection confidence threshold & 0.4\\
\hline
\gls{iou} threshold (\gls{nms}) & 0.3\\
\hline
Track maximum age & 30 frames\\
\hline
Track minimum hits & 3 frames\\
\hline
\gls{lstm} input sequence length & 50 timesteps\\
\hline
\gls{lstm} prediction horizon & 20 timesteps\\
\hline
Training batch size & 32\\
\hline
Training epochs & 100\\
\hline
\gls{ais} matching radius & 5.5 km\\
\hline
\gls{cpa} critical alert threshold & 50 m\\
\hline
Average input resolution & 1280$\times$720\\
\hline
Target processing speed & 30 \gls{fps}\\
\hline
\end{tabular}
\end{table}

\section{Performance and Stress Testing}

An essential dimension of the evaluation involved stress testing the software architecture under a heavy computational load. Given the objective to serve as a real-time surveillance platform, the system was subjected to dense maritime video feeds populated with dozens of simultaneously active vessels. The performance monitoring concentrated heavily on identifying latency bottlenecks that could potentially drop the application's processing capability below the critical 30 frames-per-second requirement. 

During these high-density trials, particular attention was paid to the execution times of the DeepOCSORT multi-object tracker and the LSTM forecasting modules. As the number of overlapping ship trajectories naturally increased in crowded waterways, the testing framework closely monitored CPU and GPU memory utilization alongside frame inference latency. This rigorous stress testing ultimately validated the necessity of the system's kinematic linear regression fallback mechanism, proving that it successfully prevents the pipeline from stalling while calculating initial trajectory predictions for newly identified targets.

\section{Test Cases}

A series of targeted test cases were executed to evaluate different models, architectures, and features during the system's development. The structured outcomes of these test cases are detailed below.

\begin{table}[H]
\centering
\caption{Test Case 1: Evaluation of Pre-trained YOLO Architectures}
\label{tab:testcase1}
\begin{tabular}{|p{0.3\textwidth}|p{0.6\textwidth}|}
\hline
\textbf{Sl No. of Test Case} & 1 \\
\hline
\textbf{Name of Test Case} & Evaluation of Pre-trained YOLO Architectures \\
\hline
\textbf{Feature being Tested} & Zero-shot Object Detection Efficacy and Computational Overhead \\
\hline
\textbf{Description} & Assess the baseline performance of unmodified YOLO architectures on maritime datasets. The goal is to determine their native detection precision, bounding box regression accuracy, and overall spatial memory requirements before any domain-specific fine-tuning is applied. \\
\hline
\textbf{Sample Input} & High-definition maritime surveillance video streams (1080p resolution) featuring varying sea states, weather conditions (fog, rain, glare), and dense multi-vessel scenarios. \\
\hline
\textbf{Expected Output} & High mean Average Precision (mAP) across vessel classes, coupled with a minimal memory footprint and consistent inference speeds conducive to edge deployment. \\
\hline
\textbf{Actual Output} & Demonstrated adequate baseline detection on isolated vessels without fine-tuning. However, accuracy dropped significantly in complex scenes with heavy glare, and the parameter count for heavier models vastly exceeded target hardware constraints, causing inference latency. \\
\hline
\textbf{Remarks} & System requirements successfully validated. The testing conclusively proved that fine-tuning of lighter YOLOv8 variants (the 'nano' or 'small' models) is absolutely necessary. \\
\hline
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Test Case 2: Feature-based Re-identification via Database Mapping}
\label{tab:testcase2}
\begin{tabular}{|p{0.3\textwidth}|p{0.6\textwidth}|}
\hline
\textbf{Sl No. of Test Case} & 2 \\
\hline
\textbf{Name of Test Case} & Feature-based Re-identification via Database Mapping \\
\hline
\textbf{Feature being Tested} & Multi-Target Re-identification and Historical Trajectory Mapping \\
\hline
\textbf{Description} & Extract bounding box visual features via a Convolutional Neural Network (CNN) and store high-dimensional embeddings in a relational database for inter-frame vessel matching. \\
\hline
\textbf{Sample Input} & Extracted spatial coordinates and normalized visual feature vectors \\
\hline
\textbf{Expected Output} & Low-latency query execution yielding highly accurate topological tracking across sequential frames, compatible with constrained processing units. \\
\hline
\textbf{Actual Output} & Achieved high re-identification accuracy in isolated tests; however, synchronous database I/O operations induced severe latency bottlenecks when integrated with the primary YOLO inference pipeline. \\
\hline
\textbf{Remarks} & Architecture deemed computationally unfeasible for target hardware; Test failed \\
\hline
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Test Case 3: Kinematic Multi-Object Tracking utilizing DeepOCSORT}
\label{tab:testcase3}
\begin{tabular}{|p{0.3\textwidth}|p{0.6\textwidth}|}
\hline
\textbf{Sl No. of Test Case} & 3 \\
\hline
\textbf{Name of Test Case} & Kinematic Multi-Object Tracking utilizing DeepOCSORT \\
\hline
\textbf{Feature being Tested} & Inter-frame Identity Consistency and Occlusion Handling \\
\hline
\textbf{Description} & Integrate the DeepOCSORT algorithm to maintain continuous trajectory associations and preserve unique tracking IDs utilizing bounding box kinematics and Kalman filtering. \\
\hline
\textbf{Sample Input} & Time-series YOLOv8 detection coordinates and associated confidence scores \\
\hline
\textbf{Expected Output} & Persistent identity tracking with robust handling of temporary visual occlusions and minimal computational overhead. \\
\hline
\textbf{Actual Output} & Facilitated highly efficient trajectory history management on lower-tier hardware. Identity fragmentation during missed detections was subsequently mitigated via sensor fusion with AIS telemetry. \\
\hline
\textbf{Remarks} & Systematically successful; Adopted into primary pipeline \\
\hline
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Test Case 4: Offline Geographic Visualization via GeoTIFF Rendering}
\label{tab:testcase4}
\begin{tabular}{|p{0.3\textwidth}|p{0.6\textwidth}|}
\hline
\textbf{Sl No. of Test Case} & 4 \\
\hline
\textbf{Name of Test Case} & Offline Geographic Visualization via GeoTIFF Rendering \\
\hline
\textbf{Feature being Tested} & Offline Cartographic Rendering and Spatial Mapping \\
\hline
\textbf{Description} & Ingest and render high-resolution GeoTIFF files to provide localized, offline geographic context for maritime surveillance operations. \\
\hline
\textbf{Sample Input} & Standardized high-resolution cartographic GeoTIFF tiles \\
\hline
\textbf{Expected Output} & Detailed, responsive rendering of geographic contours with minimal storage overhead. \\
\hline
\textbf{Actual Output} & Successfully rendered topographical contours offline; however, individual map segments exceeded 500 MB, violating the strict memory constraints of the deployment hardware. \\
\hline
\textbf{Remarks} & Implementation rejected due to excessive spatial complexity; Test failed \\
\hline
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Test Case 5: Real-time Telemetry Visualization via AIS and MapLibre}
\label{tab:testcase5}
\begin{tabular}{|p{0.3\textwidth}|p{0.6\textwidth}|}
\hline
\textbf{Sl No. of Test Case} & 5 \\
\hline
\textbf{Name of Test Case} & Real-time Telemetry Visualization via AIS and MapLibre \\
\hline
\textbf{Feature being Tested} & Dynamic Web-Cartography and Telemetry Sensor Fusion \\
\hline
\textbf{Description} & Establish a WebSocket interface to an open-source Automatic Identification System (AIS) stream and project live vessel telemetry onto a MapLibre vector-tile map. \\
\hline
\textbf{Sample Input} & Asynchronous AIS JSON data streams and lightweight map vector tiles \\
\hline
\textbf{Expected Output} & Reliable, low-latency rendering of live geospatial data synchronized with the visual tracking pipeline. \\
\hline
\textbf{Actual Output} & WebSocket integration proved highly efficient with minimal overhead. The sole limitation observed was incomplete geographic coverage for secondary ports within the free-tier stream. \\
\hline
\textbf{Remarks} & Highly successful; Adopted as the primary mapping infrastructure \\
\hline
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Test Case 6: Predictive Trajectory Modeling utilizing Bidirectional LSTM}
\label{tab:testcase6}
\begin{tabular}{|p{0.3\textwidth}|p{0.6\textwidth}|}
\hline
\textbf{Sl No. of Test Case} & 6 \\
\hline
\textbf{Name of Test Case} & Predictive Trajectory Modeling utilizing Bidirectional LSTM \\
\hline
\textbf{Feature being Tested} & Non-linear Kinematic Forecasting and Time-Series Analysis \\
\hline
\textbf{Description} & Deploy a Bidirectional Long Short-Term Memory (BiLSTM) neural network to analyze historical position vectors and forecast future geospatial coordinates of tracked vessels. \\
\hline
\textbf{Sample Input} & Sequential tensors representing the trailing 50 kinematic states of an identified vessel \\
\hline
\textbf{Expected Output} & Accurate geospatial predictions capable of anticipating non-linear maneuvers and acute navigational deviations. \\
\hline
\textbf{Actual Output} & The BiLSTM architecture significantly minimized both Average Displacement Error (ADE) and Final Displacement Error (FDE) compared to linear regression baselines, correctly anticipating complex maneuvers. \\
\hline
\textbf{Remarks} & Systematically successful; Integration approved \\
\hline
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Test Case 7: Automated Collision Risk Categorization (CPA)}
\label{tab:testcase7}
\begin{tabular}{|p{0.3\textwidth}|p{0.6\textwidth}|}
\hline
\textbf{Sl No. of Test Case} & 7 \\
\hline
\textbf{Name of Test Case} & Automated Collision Risk Categorization (CPA) \\
\hline
\textbf{Feature being Tested} & Proactive Spatial Risk Assessment and Alert Generation \\
\hline
\textbf{Description} & Compute the Closest Point of Approach (CPA) utilizing current kinematic vectors for active vessel pairs to trigger severity-based collision alerts. \\
\hline
\textbf{Sample Input} & Real-time geospatial coordinates, heading, and scalar velocity of overlapping vessel trajectories \\
\hline
\textbf{Expected Output} & Mathematically precise categorization of trajectory intersections into Critical, High, Medium, and Low risk thresholds with zero false-negative dropouts. \\
\hline
\textbf{Actual Output} & The algorithm accurately resolved trigonometric approach distances in real-time and consistently triggered deterministic alerts conforming to the predefined spatial thresholds. \\
\hline
\textbf{Remarks} & Highly successful; Crucial safety mechanism validated \\
\hline
\end{tabular}
\end{table}

\section{Summary}
This chapter detailed the comprehensive testing framework utilized to validate the maritime surveillance platform under realistic operational conditions. By systematically executing defined test cases, ranging from unit-level algorithm checks to end-to-end integration trials, the evaluation phase rigorously confirmed the system's robustness. The hardware and software environments were strictly standardized to prevent hyperparameter drift and to ensure that all recorded metrics remained entirely reproducible.

Through aggressive stress testing and the analysis of detailed defect logs, the architecture demonstrated exceptional operational stability, detection accuracy, and real-time processing capabilities. Crucially, the integration of DeepOCSORT tracking and BiLSTM trajectory forecasting proved capable of maintaining a fluid 30 frames-per-second throughput without exceeding commercial GPU constraints. Ultimately, these testing outcomes substantiate the platform's readiness for live deployment in complex, high-traffic maritime environments.
