\chapter{Software Requirement Specification of Maritime Surveillance and Trajectory Prediction}

This chapter outlines the formal software requirements for the proposed maritime surveillance system. It details the product perspective, core functional and performance requirements, as well as the necessary software and hardware infrastructure constraints governing the system's design and deployment.

\section{Overall Description}

\subsection{Product Perspective}
The proposed architectural solution operates as an integrated, multi-stage pipeline specifically engineered for continuous, real-time maritime monitoring. At the core of the pipeline is a structured data ingestion module that handles raw visual inputs from port surveillance systems or coastal cameras. Upon receiving the video stream, each frame is processed individually by the \gls{yolov8} object detection framework. This deep learning model is responsible for isolating vessels within the frame and generating precise bounding boxes. Once a vessel is detected, the spatial coordinates are immediately passed down to the \gls{deepocsort} tracking module. The tracking system then combines appearance-based \gls{reid} embeddings with kinematic estimations via a Kalman filter to maintain a persistent identity for every ship across consecutive frames, even in crowded or partially occluded scenarios.

As the target vessels navigate through the field of view, the system continuously logs their centroid positions. By evaluating the frame-to-frame displacement of these coordinates, critical motion dynamics such as relative speed and directional heading are computed. This historical trajectory information forms the foundational dataset for the prediction stage. The sequential coordinate data is routed into an \gls{lstm} neural network, which leverages its recurrent architecture to anticipate the vessel's subsequent movements over a defined temporal window. Ultimately, the insights generated by the detection, tracking, and prediction modules are consolidated. This aggregated data is visually superimposed onto the live video feed and plotted geospatially via a Folium-based interface, equipping maritime operators with a unified and highly intuitive situational awareness dashboard.

\subsection{Product Functions}
The primary functionalities that the system aims to deliver include:
\begin{enumerate}
\item \textbf{Vessel Identification and Tracking:} To construct a robust computer vision pipeline utilizing \gls{yolov8} for accurate ship recognition and \gls{deepocsort} for continuous multi-object tracking. This ensures dependable monitoring capabilities even when confronted with typical maritime environmental hurdles, such as reflections on the water surface and temporary visual obstructions.
\item \textbf{Trajectory Analysis and Forecasting:} To derive meaningful analytical movement data from the tracked subjects. By employing an \gls{lstm} neural network in conjunction with recursive state estimation algorithms, the system functions to predict the future navigational paths of vessels, thereby facilitating preemptive threat assessment and traffic management.
\item \textbf{Integrated Visualization and Mapping:} To engineer a holistic graphical interface that enriches the raw video feed with analytical overlays, including bounding boxes, unique identification markers, velocity vectors, and anticipated routes. Concurrently, the system maps this spatial data onto an interactive geographic layout to offer a broader perspective of the monitored maritime domain.
\end{enumerate}

\subsection{User Characteristics}
The intended users of this system are primarily maritime traffic controllers, port security personnel, and coast guard operators. These individuals generally possess a strong understanding of maritime operations and navigational rules but may not necessarily have deep technical expertise in machine learning or computer vision. Therefore, the user interface is designed to be highly intuitive, presenting complex analytical outputs, such as predicted trajectories and tracking IDs, in a clear, visually accessible manner. 

\subsection{Assumptions and Dependencies}
The successful deployment and operation of the system are contingent upon several key assumptions. The primary input is assumed to be a stable video feed originating from a stationary camera setup, which is standard for coastal and harbor surveillance infrastructure. It is expected that the target vessels will have adequate visual contrast against the surrounding water to enable effective detection. Furthermore, the foundational \gls{yolov8} model is presumed to be trained on a varied and comprehensive dataset that adequately represents the types of vessels typically encountered in the deployment area. The trajectory forecasting mechanism assumes that vessel movements follow relatively predictable and smooth dynamics, as highly erratic or sudden maneuvers can compromise the predictive accuracy of the \gls{lstm} module. On the hardware front, achieving real-time processing speeds strictly depends on the availability of a dedicated GPU within the deployment environment.

\subsection{Constraints}
The project scope is defined by certain operational and architectural limitations. At its current iteration, the system processes exclusively optical video streams and does not integrate supplementary data sources such as marine radar or live \gls{ais} broadcasts. Consequently, the effectiveness of the detection and tracking algorithms is heavily reliant on visual clarity and will naturally degrade during adverse weather events, such as dense fog, heavy rainfall, or poor lighting conditions. The accuracy of the \gls{lstm}-based prediction model is intrinsically linked to the duration and stability of the recorded historical track; fragmented or exceedingly brief tracks will result in less reliable forecasts. The current framework is optimized for analyzing a single camera feed independently and does not possess the capability to perform spatial hand-offs or stitch tracking data across multiple overlapping camera views. Additionally, the precision of the geospatial mapping feature is strictly dependent on accurate camera calibration, which is required to correctly translate pixel-based coordinates into real-world geographic coordinates.

\section{Specific Requirements}

\subsection{Functional Requirements}
The functional requirements outline the specific behaviors and capabilities the system must exhibit:
\begin{itemize}
    \item \textbf{FR1 - Object Detection:} The application shall ingest video frames and utilize the YOLOv8 architecture to identify maritime vessels, returning accurate bounding box coordinates for each detected entity.
    \item \textbf{FR2 - Multi-Object Tracking:} The system shall implement the DeepOCSORT algorithm to assign and maintain unique, persistent identifiers for all detected vessels across sequential video frames, handling temporary occlusions gracefully.
    \item \textbf{FR3 - Motion Analysis:} The software must calculate the speed and heading of each tracked vessel based on the frame-by-frame displacement of its centroid.
    \item \textbf{FR4 - Path Prediction:} The application shall employ an LSTM neural network to process historical trajectory data and forecast the future coordinates of the vessels.
    \item \textbf{FR5 - Visual Overlay:} The system must superimpose the detection bounding boxes, tracking IDs, motion vectors, and predicted paths directly onto the output video feed.
    \item \textbf{FR6 - Geospatial Mapping:} The software shall translate the on-screen pixel coordinates into geographic coordinates and plot the vessel locations and predicted paths on an interactive Folium map.
\end{itemize}

\subsection{Performance Requirements}
\begin{itemize}
    \item \textbf{PR1 - Real-Time Processing:} The integrated pipeline must process the incoming video stream at a rate that matches or closely approximates the native frame rate of the camera, ensuring real-time or near real-time situational awareness.
    \item \textbf{PR2 - Latency:} The end-to-end latency from frame ingestion to the display of the analytical overlays should not exceed a threshold that would render the information operationally obsolete (e.g., typically under 500 milliseconds per frame).
    \item \textbf{PR3 - Accuracy:} The object detection and tracking modules should maintain a high degree of precision and recall to minimize false positives and identity switches, even in moderately crowded port environments.
\end{itemize}

\subsection{Software Requirements}
The development and execution of the system rely on the following software stack:
\begin{itemize}
    \item \textbf{Programming Language:} Python 3.8 or higher.
    \item \textbf{Deep Learning Framework:} PyTorch for running the YOLOv8 and LSTM models.
    \item \textbf{Computer Vision Library:} OpenCV for image processing, frame manipulation, and overlay generation.
    \item \textbf{Parallel Computing Platform:} CUDA toolkit and cuDNN for GPU-accelerated tensor operations.
    \item \textbf{Mapping Interface:} Folium for rendering the interactive geospatial maps.
\end{itemize}

\subsection{Hardware Requirements}
To satisfy the performance requirements, particularly real-time inference, the deployment hardware must meet the following specifications:
\begin{itemize}
    \item \textbf{Processor:} A modern multi-core CPU (e.g., Intel Core i7 or AMD Ryzen equivalent).
    \item \textbf{Memory:} Minimum of 16 GB RAM to handle the video buffering and model weights.
    \item \textbf{Graphics Processing Unit:} A dedicated NVIDIA GPU with sufficient VRAM (minimum 8 GB recommended, such as an RTX 3060 or higher) with CUDA support is strictly required to accelerate the neural network inferences.
\end{itemize}

\subsection{Design Constraints}
The system's design is constrained by the inherent limitations of optical sensors. The reliance on visual data means that the system's efficacy is directly tied to the visibility conditions of the deployment area. Severe meteorological phenomena that obscure the camera's line of sight will inevitably lead to a degradation in detection confidence and tracking stability. Furthermore, the requirement for a stationary camera setup restricts the system's deployability on moving platforms, such as patrol boats or drones, without significant modifications to incorporate dynamic background subtraction and ego-motion compensation.

In addition to optical limitations, the architecture must operate under strict computational and bandwidth constraints. Real-time video processing combined with deep learning inference imposes heavy demands on local hardware, severely limiting the system's deployment footprint. To prevent thermal throttling in typical maritime control environments, the hardware must be equipped with specialized cooling solutions, increasing overall deployment costs.

Data storage and network latency present further structural boundaries. The continuous buffering of high-definition video frames alongside expanding historical trajectory databases requires rigorous data lifecycle management to prevent memory overflow. Moreover, any potential integration with cloud-based analytics or centralized command centers will be constrained by the available network bandwidth at coastal outposts, which often lack high-speed fiber connectivity. These bandwidth limitations necessitate that all primary inference processing occurs locally at the edge, placing a hard cap on the size and complexity of the neural network models that can be realistically deployed without sacrificing the crucial real-time processing metric.

\section{Summary}
This chapter has detailed the comprehensive Software Requirement Specification for the maritime surveillance project. It established the overall product perspective, outlining the modular pipeline from detection to geospatial mapping. The core functionalities, user characteristics, and the underlying assumptions governing the system's operation were thoroughly discussed. Furthermore, the chapter delineated the specific functional and performance requirements, along with the necessary software and hardware infrastructure, thereby providing a clear blueprint for the subsequent design and implementation phases.
